Spike-and-Slab Generalized Additive Models and Fast Algorithms for High-Dimensional Data

### 2021-09-07 Manuscript/02-Method.Rmd
Finish Revising Section 2.2

Following the GLM notation introduced in Section \ref{sec:intro}, we have a generalized additive model and its matrix form,
\begin{equation}\label{eq:gam}
E(Y|\bs x) = g^{-1}(a + \sum\limits^p_{j=1}B_j(x_j)) = g^{-1}(a + \sum\limits^p_{j=1} \bs \beta_j^T \bs X_j),
\end{equation}
with smoothing functions $B_j(x_j)$ of the variable $x_j, j = 1, \dots, p.$ The basis function matrix, i.e. the design matrix derived from the smoothing function $B_j(x_j)$, is denoted $\bs X_j$ for the variable $x_j$. The dimension of the design matrix depends on the choice of the smoothing function. For example, a cubic spline function with $k$ knots produces a $k+2$ elements column vector. Its corresponding smoothness penalty matrix is denoted as $\bs S_j$. $\bs \beta_j$ denotes the basis function coefficients for the $j$th variable, and the matrix transposing operation is denoted with a superscript $^T$. To note, the proposed model can include parametric forms of variables in the model, treating the parametric function as a special case of the smoothing function, e.g. $B_j(x_j) = x_j$ with the smoothness penalty matrix defined as $\bs S_j = \begin{bmatrix}0\end{bmatrix}$.

To encourage proper smoothness of the smoothing functions, we adopt the idea of smoothing penalties from smoothing spline models. The basic idea is to set up a smoothing penalty, depending on the smoothness penalty matrix via Equation \ref{eq:smoothpen}, in the posterior density function, or least squared and likelihood function in the penalized framework. However, the direct integration of smoothness penalties into sparsity penalties is not obvious. Given the smoothness penalty matrix is most often symmetric and positive semi-definite ^[ we limit the discussion to univariate smoothing function. However, the penalty matrix could possibly contain multiple matrices, e.g. tensor product smoother.], it can be integrated into the design matrix $\bs X$ via a re-parameterization. The re-parameterization follows Marra and  Wood [@Marra2011], in which eigen-decomposed the penalty matrix $\bs S = \bs U \bs D \bs U^T$, where the matrix $\bs D$ is diagonal with the eigenvalues arranged in the ascending order. To note, $\bs D$ can contain $0$ elements on the diagonal, associated with the null space of the smooth function. We further write the orthonormal matrix $\bs U$, containing the eigenvectors as column, as an (row) augmentation of two sub-matrices $\bs U \equiv \begin{bmatrix}\bs U^0 : \bs U^{\text{pen}}\end{bmatrix}$, where $\bs U^{\text{pen}}$ contains the eigenvectors (as columns) for the non-zero eigenvalues, i.e. the penalized space, and $\bs U^0$ contains the eigenvectors for the null space. We multiply the basis function matrix $\bs X$ with the orthonormal matrix $\bs U$ for the new design matrix $\bs X^\ast = \bs X \bs U \equiv \begin{bmatrix} \bs X^0 : \bs X^{\text{pen}} \end{bmatrix}$. An additional scaling step is imposed on $\bs X^{\text{pen}}$ by the non-zero eigenvalues of $\bs D$ such that the new basis function matrix $\bs X^\ast$ is scale-invariant ^[ I know this work for tensor spline (see [@Wood2006]), but I am not sure if it can be generalized to all smoothing functions]. For notation convenience, we denote $\bs X_j \equiv \begin{bmatrix} \bs X_j^0 : \bs X_j^{\text{pen}} \end{bmatrix}$ as the after re-parameterized basis function matrix for the $j$th variable hereafter. Similarly, a spline function can be expressed in the matrix form
$$
f_j(x_j) = f_j^0(x_j) + f_j^\tp (x_j) = \bs {\beta_j^0}^T \bs X_j^0 + \bs {\beta_j^\tp}^T \bs X_j^\tp ,
$$
and the generalized additive model in Equation (\ref{eq:gam}) be
\begin{equation}\label{eq:gam-repa}
E(Y|\bs x) =  g^{-1}(a + \sum\limits^p_{j=1} \bs \beta_j^T \bs X_j) = g^{-1}\left[a + \sum\limits^p_{j=1} ({\beta_j^0}^T \bs X_j^0 + {\beta_j^\tp}^T \bs X_j^\tp)\right].
\end{equation}

The re-parameterization sets up the foundation of the proposed model, which has three-fold benefits. First of all, as previously mentioned, the re-parameterization integrated the smoothness penalty into the design matrix, and hence the penalty can be expressed in the posterior density function implicitly in addition to the sparse penalty for variable and function selection. Secondly, the eigendecomposition of the smooth penalty isolates the linear space of a smooth function from the nonlinear space, improving the feasibility of bi-level variable/function selection and estimation inference. The eigendecomposition also introduced the linearly independent column spaces of the smooth function matrix, which allows imposing independent priors on the coefficients. This reduces the computational complexity that multivariate priors induces, and greatly broadens the choices of priors and further model choices. Last but not least, the scaling step of $\bs X^\tp$ further simplifies the choice of penalty priors, from column-specific penalties to a unified penalty. This step is also necessary for the proposed spike-and-slab spline prior. Conventionally, the spike-and-slab priors require scaling of predictors ^[TODO: insert a citation.]

@@

The family of spike-and-slab (SS) priors regression models dominates Bayesian high-dimensional analysis for its utility in outcome prediction and variable selection. We defer to Bai et al. [@Bai2021Review] for an in-depth introduction to spike-and-slab priors. To summary, spike-and-slab priors are a family of mixture distributions that comprises a skinny spike density $f_{spike}(\cdot)$ for weak signals and a flat slab density $f_{slab}(\cdot)$ for large signals, mathematically
$$
 \beta|\gamma \sim (1-\gamma)f_{spike}(\beta) + \gamma f_{slab}(\beta).
$$
The most distinct feature of the SS priors is that it is conditioned on a latent binary variable $\gamma$ that indicates whether the variable $x$ is included in the model. There are various spike-and-slab priors depending on the choice for the spike density $f_{spike}(\cdot)$^[TODO: should I change to notation of $f$ to not confound with the smooth functions?] and the slab density $f_{slab}(\cdot)$, see George and McCulloch [@George1993; @George1997]; Chipman [@Chipman1996] for grouped variables; Brown et al. [@Brown1998] for multivariate outcomes; Ishwaran and Rao [@Ishwaran2005]; Clyde and George [@Clyde2004] and reference therein.

@@

The major criticism of early spike-and-slab priors receives is being computationally prohibitive. [@Bai2021Review] Since then, many studies focus on alleviating the computational burden that sampling algorithms bear, which include EMVS [@Rockova2014a] and Spike-and-slab Lasso [@Rockova2018b; @Rockova2018]. Particularly, the development of Spike-and-slab Lasso  substantially improves the scalability of SS models, setting up the theoretical foundation for generalized models in -omics data analysis [@Tang2017a; @Tang2017; @Tang2018; @Tang2019]. The spike-and-slab lasso prior is composed of two double exponential distributions with mean 0 and different dispersion parameters, $0 < s_0 < s_1$, mathematically,
\begin{equation} \label{eq:ssl}
\beta | \gamma \sim (1-\gamma)DE(0, s_0) + \gamma DE(0, s_1), 0 < s_0 < s_1.
\end{equation}
The SSL also mitigated the problem of EMVS where the weak signals are not shrink to zero, favored in high-dimensional data analysis. We noticed that Bai [@Bai2021] introduces a spike-and-slab group prior in the GAM framework, where the densities of the spike and slab components take the group lasso density [@Xu2015]. Bai [@Bai2021] takes an "all-in-all-out" strategy for function selection.

### Spike-and-Slab Lasso Prior
We introduce a novel prior for GAMs, particularly for high-dimensional nonlinear effect modeling with bi-level selection. The proposed prior extends from the spike-and-slab lasso prior described in Equation \ref{eq:ssl}. Given the reparameterized deign matrix $\bs X_j = \begin{bmatrix} \bs X_j^0 : \bs X_j^\tp \end{bmatrix}^T$ for the $j$th variable, we impose a two-part SSL prior to the coefficients $\bs \beta_j$. Specifically, we impose independent group priors on the null space coefficients and on the penalized space coefficients respectively,

@@

where $\gamma^0_{j}\in\{0,1\}$ and $\gamma^\tp_{j}\in \{0,1\}^{K_j}$ are latent indicator variables, indicating if the model includes the linear effect and the nonlinear effect of the $j$th variable respectively. ; $s_0$ and $s_1$ are scale parameters, assuming $0 < s_0 < s_1$ and given. These scale parameters $s_0$ and $s_1$ can be treated as tuning parameters and optimized via cross-validation. A discussion of how to choose the scale parameters comes in Section \ref{sec:tune}. To note, this prior differs from previous group spike-and-slab lasso priors [@Tang2018; @Tang2019], where the $\bs \beta^0_j$ and $\bs \beta^\tp_j$ shares the same indicator variables $\gamma_j^0$, $\gamma_j^\tp$ respectively. It is possible to relax the assumption such that each coefficient $\beta$ to have its own latent indicator $\gamma$, but at the cost ofcomplicating the bi-level function selection. Conversely, it is possible to add a more restrictive assumption on the priors, assuming that one indicator variable decides the inclusion of both the linear effect and nonlinear effect, i.e. $\gamma_j = \gamma^0_j = \gamma^\tp_j$. This converges^[Is "becomes" better?] to the SB-GAM [@Bai2021].

The reparameterization introduced in Section \ref{sec:BHAM} grants the validity of the proposed prior. First of all, the smooth function bases are linear dependent to each other based on the definitions, and ask for extra attention. The eigeondecomposition remedies the problem. Hence our prior can be set to be conditionally independent. Secondly, the eigenvalue scaling provides a panacea to allow unified scale parameters for all bases of all smoothing functions of variables, where scaling is mandate.

The rest of the hierarchical prior follows the traditional spike-and-slab lasso piror: we set up the hyper-prior of $\bs \gamma$, to allow local adaption of the shrinkage, using a bernoulli distribution. The two indicators of the $j$th predictor, $\gamma^{0}_j$ and $\gamma^\tp_j$, shares the same probability parameter,
$$
\begin{aligned}
\gamma_{j}^{0} | \theta_j &\sim Bin(\gamma^{0}_{j}|1, \theta_j)\\
\gamma_{j}^\tp | \theta_j &\sim Bin(\gamma^\tp_{j}|1, \theta_j).
\end{aligned}
$$

This is to leverage the fact that the probability of selecting the bases of a smooth function should be similar, while allowing different penalty on the null space and penalty space of the spline function. The hyper prior of $\gamma_{j}^{0}$ decides the sparsity of the model at the function selection level, while that of $\gamma_{j}^\tp$ decides the smoothness of the spline function at basis function level. Meanwhile, we specify that $\gamma_{j}^0$ and $\gamma_{j}^\tp$ are independently distributed for analytic simplicity. We further specify the parameter $\theta_j$ follows a beta distribution with given shape parameters $a$ and $b$,
$$
\theta_j \sim Beta(a, b).
$$
The beta distribution is a conjugate prior for the binomial distribution and hence provides some computation convenience. For simplicity, we focus on a special case of beta distribution, uniform (0,1). When the variable have large effects in any of the bases, the parameter $\theta_j$ will be estimated large, which in turn encourages the model to include the rest of bases, achieving the local adaption among spline bases. Hereafter, we refer Bayesian hierarchical generalized additive models with the spike-and-slab lasso prior as the BHAM-SSL, and visually presented in Figure \ref{fig:SSprior}. \\

@@

With the re-parameterization step of the basis function matrix $X$, it is possible to generalized the SSL prior to other priors, for example normal priors for ridge-type regularization and mixture normal prior for spike-and-slab
regularization. These priors would work better in low and medium dimensional settings where the sparse assumption is not necessary. Here we focus on the mixture normal prior as a demonstration of applying continuous spike-and-slab prior in additive models. ^[TODO: enter a sentence introducing EMVS and spike-and-slab mixture normal prior].

@@

Alternative to the sampling algorithms for fitting Bayesian models, optimization algorithms focus on the maximum a posteriori (MAP) estimates and speed up the process at the cost of uncertainty measures. The earliest work for fitting spike-and-slab models using optimization algorithm is EMVS [@Rockova2014a]. Ro{\v{c}}kov{\'{a}} and George[@Rockova2014a] proposed an expectation-maximization (EM) based algorithm to fit models that use continuous mixture normal priors. 
<!-- The coefficient $\beta_j$ independently follows a mixture normal distribution $N(0, \sigma^2((1-\gamma_i)v_0 + \gamma_i v_1)$ for $0 \leq v_0 < v_1$. The parameter $\sigma^2$ follows a inverse gamma prior, $IG(v/2, v\lambda /2)$. Like in the spike-and slab family, the latent binary variables $\bs \gamma$ follows a Bernoulli prior with a hyper prior $\theta$ follows a beta distribution, or simpler the uniform (0,1).  -->
In the E step, the latent variables $\gamma$ are treated as the missing data, and calculate the posterior mean of the density function; in the M step, a ridge estimator was used to update the coefficients, and $\sigma^2, \theta$ are updated accordingly. 
<!-- As the ridge regression creates small but nonzero coefficients, a further step is needed to select the variable: EMVS further suggests to threshold the inclusion probability for variable selection or generate regularization plot with different values of the spike scale parameter. EMVS is also compatible with structured prior when dealing with grouped variables. The group structure can be imposed on the prior of the latent binary variable $\gamma$ via either logistic regression or markov random field.  -->
The same authors [@Rockova2018b; @Rockova2018] further combined the EM algorithm with coordinate descent (CD) algorithm for fitting SSL models.
<!-- the continuous spike-and-slab Gaussian prior to the spike-and-slab double exponential distribution, as known as spike-and-slab lasso (SSL). The authors also lay the theoretical discussion of comparing spike-and-slab from the penalty perspective and the Bayesian perspective. The SSL also mitigated the problem of EMVS that the weak signals are not shrink to zero. -->
Yi and his group incorporated the SSL prior in the EM-Iterative Weighted Least Square(IWLS) algorithm and proposed another EM algorithm, EM-cyclic coordinate descent, based on the Lasso penalty SSL exhibit. Such applications were used in generalized linear models[@Tang2017a], cox models [@Tang2017] and their grouped counterparts[@Tang2018; @Tang2019]. Both EM based algorithms provide deterministic solutions, which becomes a more popular property as reproducible results. Bai et al. [@Bai2020] provides an up-to-date summary of SSL methods and theoretical discussion.

>>>>

Following the GLM notation introduced in Section \ref{sec:intro}, we have a generalized additive model and its matrix form,
\begin{equation}\label{eq:gam}
E(Y|\bs x) = g^{-1}(a + \sum\limits^p_{j=1}B_j(x_j)) = g^{-1}(a + \sum\limits^p_{j=1} \bs \beta_j^T \bs X_j),
\end{equation}
with smoothing functions $B_j(x_j)$ of the variable $x_j, j = 1, \dots, p.$ The basis function matrix, i.e. the design matrix derived from the smoothing function $B_j(x_j)$, is denoted $\bs X_j$ for the variable $x_j$. The dimension of the design matrix depends on the choice of the smoothing function. For example, a cubic spline function with $k$ knots produces a $k+3$ elements column vector. Its corresponding smoothing penalty matrix is denoted as $\bs S_j$. $\bs \beta_j$ denotes the basis function coefficients for the $j$th variable, and the matrix transposing operation is denoted with a superscript $^T$. To note, the proposed model can include parametric forms of variables in the model, treating the parametric function as a special case of the smoothing function, e.g. $B_j(x_j) = x_j$ with the smoothing penalty matrix defined as $\bs S_j = \begin{bmatrix}0\end{bmatrix}$.

To encourage proper smoothing of the functions, we adopt the idea of smoothing penalties from smoothing spline models. The basic idea is to set up a smoothing penalty, described in Equation (\ref{eq:smoothpen}), in the prior density function, or least squared and likelihood function in the penalized framework. However, the direct integration of smoothing penalties into sparsity penalties is not obvious. Marra and  Wood [@Marra2011] proposed a re-parameterization procedure to accommodate the smoothing penalty implicitly. Given the smoothing penalty matrix is most often symmetric and positive semi-definite for univariate smoothing functions, we apply eigendecomposition on the penalty matrix $\bs S = \bs U \bs D \bs U^T$ , where the matrix $\bs D$ is diagonal with the eigenvalues arranged in the ascending order. To note, $\bs D$ can contain $0$ elements on the diagonal, associated with the null space of the smoothing function. We further write the orthonormal matrix $\bs U \equiv \begin{bmatrix}\bs U^0 : \bs U^{\text{pen}}\end{bmatrix}$ containing the eigenvectors as columns in cthe orresponding order. That is $\bs U^0$ contains the eigenvectors  with zero eigenvalues for the null space and $\bs U^{\text{pen}}$ contains the eigenvectors (as columns) for the non-zero eigenvalues, i.e. the penalized space. We multiply the basis function matrix $\bs X$ with the orthonormal matrix $\bs U$ for the new design matrix $\bs X^\ast = \bs X \bs U \equiv \begin{bmatrix} \bs X^0 : \bs X^{\text{pen}} \end{bmatrix}$. An additional scaling step is imposed on $\bs X^{\text{pen}}$ by the non-zero eigenvalues of $\bs D$ such that the new basis function matrix $\bs X^\ast$ can receive uniform penalty on each of its dimensions. With slight abuse of the notation, we denote $\bs X_j \equiv \begin{bmatrix} \bs X_j^0 : \bs X_j^{\text{pen}} \end{bmatrix}$ as the basis function matrix for the $j$th variable after the re-parameterization. A spline function can be expressed in the matrix form
$$
B_j(x_j) = B_j^0(x_j) + B_j^\tp (x_j) = \bs {\beta_j^0}^T \bs X_j^0 + \bs {\beta_j^\tp}^T \bs X_j^\tp ,
$$
and the generalized additive model in Equation (\ref{eq:gam}) be
\begin{equation}\label{eq:gam-repa}
E(Y|\bs x) =  g^{-1}(a + \sum\limits^p_{j=1} \bs \beta_j^T \bs X_j) = g^{-1}\left[a + \sum\limits^p_{j=1} (\bs {\beta_j^0}^T \bs X_j^0 + {\bs \beta_j^\tp}^T \bs X_j^\tp)\right].
\end{equation}

The re-parameterization step sets up the foundation of the proposed model and provides three-fold benefits. First of all, the re-parameterization integrates the smoothing penalty into the design matrix, and encourages models to properly smooth the nonlinear function in addition to the sparse penalty for function selection. Secondly, the eigendecomposition of the smoothing penalty allows the isolation of the linear from the nonlinear space, improving the feasibility of bi-level function selection and inference. The eigendecomposition facilitates the construction of orthonormal design matrix, which makes imposing independent priors on the coefficients possible. This reduces the computational complexity compared to using a multivariate priors, and greatly broadens the choices of priors and further model choices. Last but not least, the scaling step of $\bs X^\tp$ further simplifies the choice of priors, from column-specific priors to a unified prior.

@@

The family of spike-and-slab (SS) priors regression models dominates Bayesian high-dimensional analysis for its utility in outcome prediction and variable selection. We defer to Bai et al. [@Bai2021Review] for an in-depth introduction to spike-and-slab priors. To summarize, spike-and-slab priors are a family of mixture distributions that comprises a skinny spike density $f_{\text{spike}}(\cdot)$ for weak signals and a flat slab density $f_{\text{slab}}(\cdot)$ for strong signals, mathematically
$$
 \beta|\gamma \sim (1-\gamma)f_{\text{spike}}(\beta) + \gamma f_{\text{slab}}(\beta).
$$
The most distinct feature of SS priors is that it is conditioned on a latent binary variable $\gamma$ that indicates whether the variable $x$ is included in the model. There are various spike-and-slab priors depending on the choice for the spike density $f_{\text{spike}}(\cdot)$ and the slab density $f_{\text{slab}}(\cdot)$, see George and McCulloch [@George1993; @George1997]; Chipman [@Chipman1996] for grouped variables; Brown et al. [@Brown1998] for multivariate outcomes; Ishwaran and Rao [@Ishwaran2005]; Clyde and George [@Clyde2004] and reference therein.

@@

The major criticism of early spike-and-slab priors receive is being computationally prohibitive. [@Bai2021Review] Since then, many studies focus on alleviating the computational burden that sampling algorithms bear, which include EMVS [@Rockova2014a] and Spike-and-slab Lasso [@Rockova2018b; @Rockova2018]. Particularly, the development of Spike-and-slab Lasso  substantially improves the scalability of SS models, setting up the theoretical foundation for generalized models in -omics data analysis [@Tang2017a; @Tang2017; @Tang2018; @Tang2019]. The spike-and-slab lasso prior is composed of two double exponential distributions with mean 0 and different dispersion parameters, $0 < s_0 < s_1$, mathematically,
\begin{equation} \label{eq:ssl}
\beta | \gamma \sim (1-\gamma)DE(0, s_0) + \gamma DE(0, s_1), 0 < s_0 < s_1.
\end{equation}
The SSL also mitigates the problem of EMVS where the weak signals are not shrink to zero, and hence is prefered in high-dimensional data analysis. We notice that Bai [@Bai2021] is the first to apply spike-and-slab lasso prior in the GAM framework, where the densities of the spike and slab components take the group lasso density [@Xu2015]. Bai [@Bai2021] takes an "all-in-all-out" strategy for function selection.

### Spike-and-Slab Lasso Prior
We introduce a novel prior for GAMs, particularly for high-dimensional nonlinear modeling with bi-level selection. The proposed prior extends from the spike-and-slab lasso prior described in Equation (\ref{eq:ssl}). Given the re-parameterized deign matrix $\bs X_j = \begin{bmatrix} \bs X_j^0 : \bs X_j^\tp \end{bmatrix}$ for the $j$th variable, we impose a two-part SSL prior to the coefficients $\bs \beta_j$. Specifically, we impose independent group priors on the null space coefficients and on the penalized space coefficients respectively,

@@

where $\gamma^0_{j}\in\{0,1\}$ and $\gamma^\tp_{j}\in \{0,1\}^{K_j}$ are latent indicator variable and vectors, indicating if the model includes the linear effect and the nonlinear effect of the $j$th variable respectively. $s_0$ and $s_1$ are scale parameters, assuming $0 < s_0 < s_1$ and given. These scale parameters $s_0$ and $s_1$ can be treated as tuning parameters and optimized via cross-validation. A discussion of how to choose the scale parameters comes in Section \ref{sec:tune}. To note, this prior differs from previous group spike-and-slab lasso priors [@Tang2018; @Tang2019], where the $\bs \beta^0_j$ and $\bs \beta^\tp_j$ shares the same indicator variables $\gamma_j^0$, $\gamma_j^\tp$ respectively. It is possible to add a more restrictive assumption on the priors, assuming that one indicator variable decides the inclusion of both the linear effect and nonlinear effect, i.e. $\gamma_j = \gamma^0_j = \gamma^\tp_j$.  This converges to the SB-GAM [@Bai2021]. Conversely, it is also possible to relax the assumption such that each coefficient $\beta$ to have its own latent indicator $\gamma$, but at the cost of complicating the bi-level function selection. This reduces the proposed prior to the classic spike-and-slab lasso prior.

The re-parameterization introduced in Section \ref{sec:BHAM} grants the validity of the proposed prior. First of all, the smoothing function bases are linear dependent and ask for extra attention. The eigeondecomposition remedies the problem and hence our prior can be set to be conditionally independent. Secondly, the eigenvalue scaling provides a panacea to allow unified scale parameters for all bases of all smoothing functions of variables.

The rest of the hierarchical prior follows the traditional spike-and-slab lasso prior: we set up hyper-priors of $\bs \gamma$ to allow local adaption of the shrinkage using a Bernoulli distribution (binomial distribution of one trial). The two indicators of the $j$th predictor, $\gamma^{0}_j$ and $\gamma^\tp_j$, shares the same probability parameter $\theta_j$,
$$
\begin{aligned}
\gamma_{j}^{0} | \theta_j &\sim Bin(\gamma^{0}_{j}|1, \theta_j)\\
\gamma_{j}^\tp | \theta_j &\sim Bin(\gamma^\tp_{j}|1, \theta_j).
\end{aligned}
$$
This is to leverage the fact that the probability of selecting the bases of a smoothing function should be similar, while allowing different penalty on the null space and penalty space of the smoothing function. The hyper prior of $\gamma_{j}^{0}$ decides the sparsity of the model at the function selection level, while that of $\gamma_{j}^\tp$ decides the smoothness of the spline function at basis function level. Meanwhile, we specify that $\gamma_{j}^0$ and $\gamma_{j}^\tp$ are independently distributed for analytic simplicity. We further specify the parameter $\theta_j$ follows a beta distribution with given shape parameters $a$ and $b$,
$$
\theta_j \sim Beta(a, b).
$$
The beta distribution is a conjugate prior for the binomial distribution and hence provides some computation convenience. For simplicity, we focus on a special case of beta distribution, uniform (0,1). When the variable have large effects in any of the bases, the parameter $\theta_j$ will be estimated large, which in turn encourages the model to include the rest of bases, achieving the local adaption among spline bases. Hereafter, we refer Bayesian hierarchical generalized additive models with the spike-and-slab lasso prior as the BHAM-SSL, and visually presented in Figure \ref{fig:SSprior}. \\

@@

With the re-parameterization step of the basis function matrix $X$, it is possible to generalized the SSL prior to other priors, for example normal priors for ridge-type regularization and mixture normal prior for spike-and-slab regularization. These priors would work better in low and medium dimensional settings where the sparse assumption is not necessary. Here we focus on the mixture normal prior as a demonstration of applying continuous spike-and-slab prior in additive models.

@@

Alternative to the sampling algorithms for fitting Bayesian models, optimization algorithms focus on the maximum a posteriori (MAP) estimates and speed up the process at the cost of uncertainty measures. The earliest work for fitting spike-and-slab models using optimization algorithm is EMVS [@Rockova2014a]. Rockova and George[@Rockova2014a] proposed an expectation-maximization (EM) based algorithm to fit models that use continuous mixture normal priors. 
<!-- The coefficient $\beta_j$ independently follows a mixture normal distribution $N(0, \sigma^2((1-\gamma_i)v_0 + \gamma_i v_1)$ for $0 \leq v_0 < v_1$. The parameter $\sigma^2$ follows a inverse gamma prior, $IG(v/2, v\lambda /2)$. Like in the spike-and slab family, the latent binary variables $\bs \gamma$ follows a Bernoulli prior with a hyper prior $\theta$ follows a beta distribution, or simpler the uniform (0,1).  -->
In the E step, the latent variables $\gamma$ are treated as the missing data, and calculate the posterior mean of the density function; in the M step, a ridge estimator was used to update the coefficients, and $\sigma^2, \theta$ are updated accordingly. 
<!-- As the ridge regression creates small but nonzero coefficients, a further step is needed to select the variable: EMVS further suggests to threshold the inclusion probability for variable selection or generate regularization plot with different values of the spike scale parameter. EMVS is also compatible with structured prior when dealing with grouped variables. The group structure can be imposed on the prior of the latent binary variable $\gamma$ via either logistic regression or markov random field.  -->
The same authors [@Rockova2018b; @Rockova2018] further combined the EM algorithm with coordinate descent (CD) algorithm for fitting SSL models.
<!-- the continuous spike-and-slab Gaussian prior to the spike-and-slab double exponential distribution, as known as spike-and-slab lasso (SSL). The authors also lay the theoretical discussion of comparing spike-and-slab from the penalty perspective and the Bayesian perspective. The SSL also mitigated the problem of EMVS that the weak signals are not shrink to zero. -->
Yi and his group incorporated the SSL prior in the EM-Iterative Weighted Least Square(IWLS) algorithm and proposed another EM algorithm, EM-cyclic coordinate descent, based on the Lasso penalty SSL exhibit. Such applications were used in generalized linear models[@Tang2017a], cox models [@Tang2017] and their grouped counterparts[@Tang2018; @Tang2019]. Both EM based algorithms provide deterministic solutions, which becomes a more popular property as reproducible results. Bai et al. [@Bai2020] provides an up-to-date summary of SSL methods and theoretical discussion.
